---
title: "Chapter 9 Hierarchical Modeling"
author: "Jim Albert"
date: "2022-09-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

```{r}
library(LearnBayes)
library(readr)
```

## Estimating OGT Success Rates

Read in the OGT writing dataset.

```{r}
ogt <- read_csv("../data/writing_OGT.csv")
```

Create a data matrix, y in first column and n in second column.

```{r}
d <- cbind(ogt$Advanced, ogt$N)
```

Using the `betabinexch()` function containing the log of the posterior of (logit $\eta$, $\log K$), find normal approximation to posterior.

```{r}
fit <- laplace(betabinexch, c(-3, 4), d)
```
Output posterior mode.

```{r}
fit$mode
```

By transforming back, estimates at $\eta$ and $\log K$.

```{r}
(eta <- exp(fit$mode[1]) / (1 + exp(fit$mode[1])))
(K <- exp(fit$mode[2]))
```

Instead implement a random walk Metropolis algorithm to simulate 10,000 draws of posterior.

```{r}
sim_draws <- rwmetrop(betabinexch, 
                      proposal = list(var = fit$var,
                                      scale = 2),
                      start = c(-3, 4),
                      m = 10000, d)
```

Acceptance rate of algorithm?

```{r}
sim_draws$accept
```
Using simulated draws, find summaries of posterior of logit $\eta$ and $\log K$.

Quantiles of logit $\eta$:

```{r}
quantile(sim_draws$par[, 1],
         c(0.025, 0.25, 0.50, 0.75, 0.95))
```
Quantiles of $\log K$:

```{r}
quantile(sim_draws$par[, 2],
         c(0.025, 0.25, 0.50, 0.75, 0.95))
```
Estimates at $p_i$.  I write a short function to compute  posterior summaries for a single $p_i$.  I simulate values $p_i$ by first simulating draws of $(\eta, K)$ and then simulating values of $p_i$ from the beta posterior.

```{r}
summaries_pi <- function(j){
  invlogit <- function(y){
    exp(y) / (1 + exp(y))
  }
  eta <- invlogit(sim_draws$par[, 1])
  K <- exp(sim_draws$par[, 2])
  y <- d[, 1]
  n <- d[, 2]
  pj <- rbeta(10000, y[j] + K * eta, 
                     n[j] - y[j] + K * (1 - eta))
  c(mean(pj), sd(pj), quantile(pj, c(.05, .95)))
}
(post_moments <- t(sapply(1:9, summaries_pi)))
```

## Normal/Normal Model - Efron and Morris data

```{r}
efronmorris <- read_csv("../data/efron_morris.csv")
```
Transform proportions -- reexpressed values are approximately normal($\theta_i, 0.11$)

```{r}
logit <- function(y){log(y) - log(1 - y)}
invlogit <- function(y){exp(y) / (1 + exp(y))}
y <- logit(efronmorris$r / efronmorris$n)
d <- cbind(y,  0.11)
```

Log posterior of mean $\mu$ and $\log \tau$ available in the function `normnormexch()`.

Simulate draws of posterior of $(\mu, \log \tau)$ using random walk Metropolis.

First do normal approximation to get reasonable estimate at posterior variance.

```{r}
fit <- laplace(normnormexch, c(0, 0),
               d)
```

Then use this posterior variance estimate in the proposal distribution for the random walk Metropolis.

```{r}
sim_draws <- rwmetrop(normnormexch, 
                      proposal = list(var = fit$var,
                                      scale = 2),
                      start = c(-1, -1.5),
                      m = 10000, d)
```

Here the acceptance rate is about 34% which is reasonable.

Get estimates at $\mu$ and $\tau$ from the simulated draws.

```{r}
(mu <- mean(sim_draws$par[, 1]))
(tau <- mean(exp(sim_draws$par[, 2])))
```

Find estimate at shrinkage
$$
F(\tau^2) = \frac{1/\tau^2}{1/\tau^2 + 1/\sigma^2}
$$

```{r}
(Ft <- (1 / tau ^ 2) / (1 / tau ^ 2 + 1 / 0.11))
```
Interesting -- this shrinkage is larger than the value 0.68 quoted in the notes.

Estimates at the $\theta_i$ are found.  Then these estimates are transformed back to get estimates at the $p_i$.
 
Create a table like Table 9.4.

```{r}
theta_est <- y * (1 - Ft) + mu * Ft
p_est <- invlogit(theta_est)
cbind(efronmorris$r / efronmorris$n,
      y, theta_est, p_est)
```

Here is a graph of the posterior of $\tau$. (It seems that Figure 9.2 is really a graph of the posterior of $\tau$, not $\tau ^ 2$.)

```{r}
plot(density(exp(sim_draws$par[, 2])))
```

