---
title: "Chapter 6 Many Parameters"
author: "Jim Albert"
date: "2022-09-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

```{r}
library(LearnBayes)
library(readr)
```

## Comparing Two Poisson Means

```{r}
web_counts <- read_csv("../data/website_counts.csv")
```

Compute $n_A$, $n_B$, $s_A$ and $s_B$.

```{r}
weekend <- web_counts$Count[web_counts$Type 
                            == "Weekend"]
weekday <- web_counts$Count[web_counts$Type 
                            == "Weekday"]
na <- length(weekend)
nb <- length(weekday)
sa <- sum(weekend)
sb <- sum(weekday)
```

Plot posterior of ratio of means 
$$
\gamma = \frac{\lambda_B}{\lambda_A}
$$

Marginal posterior of $\gamma$ is given by
$$
g(\gamma | {\rm data}) \propto \frac{\gamma^{s_B+a_g-1} \exp(-b_g \gamma)}{(n_A+n_B \gamma + b_0)^{s_A + s_B + a_0}}.
$$

Set prior parameters:

```{r}
a0 <- 2
b0 <- 1
ag <- 8
bg <- 8
```

Write a function that computes the logarithm of the posterior of $\gamma$.

```{r}
logpost <- function(g){
  (sb + ag - 1) * log(g) - bg * g - 
    (sa + sb + a0) * log (na + nb * g + b0)
}
```
Graph of the posterior.

```{r}
curve(exp(logpost(x) - logpost(1.66)), 1.2, 2.2,
      ylab = "Density")
```

Approximate this posterior with a normal curve.

```{r}
fit <- laplace(logpost, 1.6)
c(fit$mode, sqrt(fit$var))
```

## Learning About a Sample Size and a Probability

### Construction of a prior

We decompose the prior of $(AB, p)$ as follows:
$$
g(AB, p) = g(AB) g(p | AB)
$$

1.  Prior on $AB$.

Using data from the 2019 baseball season, we construct a density plot of the values of $AB$ between 400 and 700.  By trial-and-error, we see that the normal(505, 90) truncated on the interval (400, 700) is a reasonable approximation to the density plot.


```{r}
library(dplyr)
library(Lahman)
Batting %>% 
  filter(yearID == 2019) %>% 
  group_by(playerID) %>% 
  summarize(AB = sum(AB)) %>% 
  filter(AB >= 400, AB <= 700) -> S
trun_normal <- function(x){
  dnorm(x, 505, 90) / (pnorm(700, 505, 90) - 
                         pnorm(400, 505, 90))
}
plot(density(S$AB),
     main = "Density Estimate of AB with Normal Approximation", 
     xlab = "K")
curve(trun_normal(x), 400, 700,
      add = TRUE, col = "red")
```

2.  Prior on $p$ conditional on $AB$

Using the same 2019 data, we construct a scatterplot of $AB$ against the batting average $AVG = H / AB$ and add a least-squares line.

```{r}
Batting %>% 
  filter(yearID == 2019) %>% 
  group_by(playerID) %>% 
  summarize(H = sum(H), AB = sum(AB)) %>% 
  filter(AB >= 400, AB <= 700) %>% 
  mutate(AVG = H / AB) -> S1
plot(S1$AB, S1$AVG)
abline(lm(AVG ~ AB, data = S1))
```

Least-squares fit:

```{r}
lm(AVG ~ AB, data = S1)$coef
```
Our prior for $p$ conditional on $AB$ will be beta($K \eta, K (1 - \eta)$) where $K = 400$ and the prior mean is 
$$
\eta = 0.199952 + 0.000135 \,AB
$$

Contour plot of the joint prior on $(AB, p)$:

```{r}
log_prior <- function(theta, d = 0){
  AB <- theta[1]
  p <- theta[2]
  eta <- 0.199952 + 0.000135 * AB
  log(trun_normal(AB)) +
    dbeta(p, 400 * eta, 400 * (1 - eta), log = TRUE)
}
mycontour(log_prior, c(400, 700, .18, .42), 0,
          xlab = "K", ylab = "p",
          main = "Contour Plot of Prior")
```

## Computation of the posterior and inference

Write a function to compute the log of the posterior.

```{r}
log_posterior <- function(theta, H){
  AB <- theta[1]
  p <- theta[2]
  eta <- 0.199952 + 0.000135 * AB
  log_prior <- log(trun_normal(AB)) +
    dbeta(p, 400 * eta, 400 * (1 - eta), log = TRUE)
  log_likelihood <- lgamma(AB + 1) - lgamma(H + 1) -
    lgamma(AB - H + 1) + H * log(p) + 
    (AB - H) * log(1 - p)
  log_prior + log_likelihood
}
```

Construct a contour plot of the posterior of $(K, p)$:

```{r}
mycontour(log_posterior, c(400, 700, .18, .42), 200,
          xlab = "K", ylab = "p",
          main = "Contour Plot of Posterior")
```

To summarize this posterior, we draw a simulated sample of parameters from the grid using the `simcontour()` function.

```{r}
sim_draws <- simcontour(log_posterior, 
                        c(400, 700, .18, .42), 
                        200, 10000)
```

Here are density estimates of the simulated draws of $K$ and $p$ -- these are the marginal posterior densities.

```{r}
plot(density(sim_draws$x),
     main = "Posterior Density of K",
     ylab = "Density",
     xlab = "K")
plot(density(sim_draws$y),
     main = "Posterior Density of p",
     ylab = "Density",
     xlab = "p")
```



