---
title: "Chapter 8 MCMC"
author: "Jim Albert"
date: "2022-09-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

## Comparing Two Poisson Means

```{r}
library(LearnBayes)
library(readr)
```


Program the exact log posterior of ($\theta, \gamma$):

```{r}
two_poisson <- function(param, dp){
  theta <- param[1]
  gamma <- param[2]
  dp$sa * log(theta) - dp$na * theta +
    dp$sb * log(theta * gamma) - 
    dp$nb * theta * gamma +
    (dp$a0 - 1) * log(theta) - dp$b0 * theta +
    (dp$ag - 1) * log(gamma) - dp$bg * gamma
}
```

Read in the data.  The list `dp` contains all of the summary stats from the data and the values of the prior parameters.

```{r}
web_counts <- read_csv("../data/website_counts.csv")
d_weekend <- web_counts$Count[web_counts$Type 
                              == "Weekend"]
d_weekday <- web_counts$Count[web_counts$Type 
                              == "Weekday"]
dp <- list(na = length(d_weekend),
             sa = sum(d_weekend),
             nb = length(d_weekday),
             sb = sum(d_weekday),
             ag = 1, a0 = 1, bg = 1, b0 = 1)
```


Use the `mycontour()` function to plot contours of the log posterior density of ($\theta, \gamma$).

```{r}
mycontour(two_poisson, c(9.5, 18, 1, 2.5), dp,
          xlab = "THETA", ylab = "GAMMA")
```

Write a function to implement Gibbs sampling.  The inputs are the starting value of $\theta$, the number of cycles `m` and the list containing the data and prior parameters.

```{r}
mygibbs <- function(theta0, m, dp){
  theta <- theta0
  output <- matrix(0, m, 2)
  for(j in 1:m){
    gamma <- rgamma(1, shape = dp$ag + dp$sb,
                       rate = dp$bg + dp$nb * theta)
    theta <- rgamma(1, shape = dp$a0 + dp$sa + dp$sb, 
                       rate = dp$b0 + 
                              dp$na + dp$nb * gamma)
    output[j, ] <- c(theta, gamma)
  }
  output
}
```

Run the Gibbs sampler for 1000 cycles starting at $\theta = 12$.

```{r}
sim_pars <- mygibbs(12, 1000, dp)
```

Redraw the contour plot of the log posterior density and display the simulated draws on top.

```{r}
mycontour(two_poisson, c(9.5, 18, 1, 2.5), dp,
          xlab = "THETA", ylab = "GAMMA")
points(sim_pars)
```


